---
title: "Exposure and Statistics Overview"
author: "Bryan Mayer"
date: "2019-04-12"
output: workflowr::wflow_html
---

Here, we calculate some of the initial (pre-model) results from the infant cohort and exposure characteristcs. 
- Demographics
- Initial survial curves
- Exposure assessment

```{r,  warning = F, message = F, echo = F}
knitr::opts_chunk$set(
  comment = NA,
  fig.align = "center",
  tidy = FALSE
)

```

```{r, echo = F, message = F, warning = F}
library(survival)
library(tidyverse)
library(conflicted)
library(kableExtra)
library(cowplot)
conflict_prefer("filter", "dplyr")

theme_set(
  theme_bw() +
    theme(panel.grid.minor = element_blank(),
          legend.position = "top")
)

source("analysis/plot_labels.R")

exposure_data = read_csv("data/exposure_data.csv")
exposure_data_long = read_csv("data/exposure_data_long.csv")

leg_plot = ggplot(data = exposure_data_long, aes(x = idpar, y = count, fill = factor(obs_infected))) +
  geom_tile() +
  scale_fill_manual("", values = infection_labels$colours, breaks = infection_labels$breaks,
                      labels = infection_labels$labels)
trans_legend = gtable::gtable_filter(ggplot_gtable(ggplot_build(leg_plot + theme(legend.position = "top"))), "guide-box")

range_str = function(x, digits = 3) paste(round(range(x), digits), collapse = " - ")
IQR_range_str = function(x, digits = 3) paste(round(quantile(x, c(0.25, 0.75)), digits), collapse = " - ")
  
```

# Demographics

## Infant ages

```{r}
exposure_data %>%
  select(FamilyID, enrollment_age) %>%
  distinct() %>%
  summarize(
        N = n(),
        enroll_median_age_days = median(enrollment_age),
        IQR = paste(quantile(enrollment_age, c(0.25, 0.75)), collapse = ", "),
        range_days = paste(range(enrollment_age), collapse = ", ")
        ) %>%
  kable() %>% kable_styling(full_width = F)

```

## Mom HIV

```{r}
exposure_data %>% select(FamilyID, momhiv) %>% 
  distinct() %>%
  group_by(momhiv) %>%
  summarize(N = n()) %>% 
  kable() %>% kable_styling(full_width = F)
```


# Survival analysis

```{r}

exposure_data %>% 
  group_by(virus, FamilyID) %>% 
  summarize(obs_infected = max(infectious_1wk),
            is_infected = max(infected)) %>% group_by(virus) %>%
  summarize(
    total_infants = n_distinct(FamilyID),
    total_infected = sum(is_infected),
    total_outcome = sum(obs_infected)
    ) %>%
  kable() %>%
  kable_styling(full_width = F)

```


```{r}
surv_data = exposure_data %>% 
  group_by(FamilyID, virus, momhiv, final_infant_wk) %>%
  summarize(
    infected = max(infected)
  )

surv_fit = surv_data %>%
  group_by(virus) %>%
  nest() %>%
  mutate(
    surv_mod = map(data, ~survfit(Surv(final_infant_wk, infected) ~ 1, data = .)),
    surv_mod_hiv = map(data, ~survfit(Surv(final_infant_wk, infected) ~ momhiv, data = .)),
    logrank = map_dbl(data, ~coin::pvalue(coin::logrank_test(Surv(final_infant_wk, infected) ~ factor(momhiv),
                                                             data = ., distribution = "exact")))
    ) %>%
  select(-data)

surv_fit %>%
  select(virus, logrank) %>%
  rename(`Mother HIV Log-rank` = logrank) %>%
  kable() %>% kable_styling(full_width = F)


surv_res = pmap_df(surv_fit, function(virus, surv_mod, surv_mod_hiv, logrank){
   broom::tidy(surv_mod) %>%
    mutate(strata = "Pooled") %>%
    bind_rows(broom::tidy(surv_mod_hiv)) %>%
    mutate(
      virus = virus,
      momhiv = str_remove_all(strata, "momhiv=")
      ) %>%
    bind_rows(crossing(virus = virus, time = -1e-12, estimate = 1, momhiv = c("Pooled", "neg", "pos")))
})

```


```{r survival-plot, warning= F}

surv_res %>%
  arrange(virus, momhiv, time) %>%
  ggplot(aes(time, estimate, colour = momhiv)) + 
  geom_step() +
  geom_point(aes(shape = n.censor > 0)) +
  scale_shape_manual(guide = F, values = c(-1, 3)) +
  scale_x_continuous("Weeks after infant birth", breaks = 0:10 * 10) +
  scale_y_continuous("Proportion uninfected", expand = c(0.01, 0)) +
  geom_vline(xintercept = 52, colour = "black", linetype = "dashed") +
  scale_color_discrete("", breaks = c("neg", "pos", "Pooled"), 
                     labels = c("Mother HIV-", "Mother HIV+", "Pooled")) +
  geom_text(data= surv_fit, aes(label = str_c("p = ", round(logrank, 2))), 
            x = Inf, y = Inf, colour = "black", vjust = 1.2, hjust = 1.2) +
  facet_wrap(~virus) +
  theme(legend.position = "top")

```

# Exposure Analysis

## Overall

```{r}

exposure_data_long %>% 
  group_by(virus, idpar) %>% 
  summarize(
    total = n(),
    total_interpolate = sum(interpolated),
    pct_interpolate = mean(interpolated) 
  ) %>%
  kable() %>%
  kable_styling(full_width = F)

```


```{r exposure-profile, fig.height=8}

exposure_data_summary = exposure_data_long %>%
  mutate(
    pos_count = count > 0
    ) %>%
  group_by(virus, FamilyID, obs_infected, idpar) %>%
  mutate(
    interpolated_pct = 100*mean(interpolated),
    total_pos = sum(pos_count),
    pct_pos = 100 * mean(pos_count)
    ) %>%
  group_by(pct_pos, total_pos, interpolated_pct, add = T) %>%
  summarise_at(vars(count), funs(N = n(), mean = mean, median = median, maximum = max))

plot_labels = exposure_data_summary %>%
  gather(stat, estimate, mean, maximum, pct_pos) %>%
  group_by(stat) %>%
  summarize(min_lim = min(estimate), max_lim = ceiling(max(estimate))) %>%
  left_join(tibble(stat = c("pct_pos", "mean", "maximum"), out_lab = c("Percent Positive", "Mean VL", "Max. VL"))) %>%
  mutate(stat = factor(stat, levels = c("pct_pos", "mean", "maximum"))) %>%
  arrange(stat) %>%
  ungroup() %>%
  mutate(letter_code = 1:3)

pl_exposure = map(plot_labels$stat %>% levels(), function(s){
    
    tmp_theme = theme(
      legend.position = "none",
      axis.title = element_text(size = 10),
      axis.text = element_text(size = 9)
    )
    
    pl_lab = subset(plot_labels, stat == s)
    out_label = pl_lab$out_lab[1]
    lower_limit = pl_lab$min_lim
    upper_limit = pl_lab$max_lim
    
    
    pl1 = exposure_data_summary %>%
      gather(stat, estimate, mean, maximum, pct_pos) %>%
      select(-median) %>%
      spread(idpar, estimate) %>%
      filter(stat == s) %>%
      ggplot(aes(x = S, y = M, colour = factor(obs_infected))) +
      geom_point() +
      geom_abline() +
      geom_point() +
      scale_colour_manual(values = c("#4393C3", "#D6604D")) +
      scale_y_continuous(paste("Mother", out_label), limits = c(lower_limit, upper_limit)) +
      scale_x_continuous(paste("Secondary Children", out_label), limits = c(lower_limit, upper_limit)) +
      facet_wrap(~virus) +
      tmp_theme
    
    pl2 = exposure_data_summary %>%
      ungroup() %>%
      mutate(
        obs_infected = factor(obs_infected),
        idpar = fct_recode(fct_rev(idpar), "Secondary\nChildren" = "S", "Mother" = "M", "Household\nSum" = "HH")
        ) %>%
      ggplot(aes_string(x = "idpar", y = s, colour = "obs_infected")) +
      geom_boxplot() +
      scale_colour_manual(values = c("#4393C3", "#D6604D")) +
      geom_point(position = position_dodge(width = 0.75)) +
      scale_y_continuous(out_label, limits = c(lower_limit, upper_limit)) +
      scale_x_discrete("") +
      facet_wrap(~virus) +
      tmp_theme +theme(axis.text.x = element_text(size = 7))
    
    plot_grid(pl1, pl2, nrow = 1, labels = LETTERS[c(pl_lab$letter_code, pl_lab$letter_code + 3)])
    
})

plot_grid(plot_grid(plotlist = pl_exposure, nrow = 3), trans_legend, nrow = 2, rel_heights = c(11, 1))

```


```{r}

exposure_data_tests = 
  exposure_data_summary %>% 
  gather(stat, estimate, mean, maximum, pct_pos) %>%
  group_by(virus, stat, idpar) %>%
  mutate(obs_infected = factor(obs_infected)) %>%
  summarize(
    wilcox_pvalue =  coin::pvalue(coin::wilcox_test(estimate ~ obs_infected, distribution = "exact"))
  )


overall_summary = exposure_data_summary %>% 
  gather(stat, estimate, mean, maximum, pct_pos) %>%
  group_by(virus, stat, idpar, obs_infected) %>%
  summarize(
    median = median(estimate)
  ) %>%
  mutate(obs_infected = recode_factor(obs_infected, `1` = "Transmission", `0` = "No transmission", .ordered = T)) %>%
  spread(obs_infected, median) %>%
  mutate(Difference = `Transmission`-`No transmission`) %>%
  left_join(exposure_data_tests)

overall_summary %>%
  kable(digits = 3) %>%
  kable_styling(full_width = F) %>%
  collapse_rows(columns = 1:2)

```


## Household composition

Household sum composition was determined and reported by taking the mean proportion over measurements within a household. The summary across the households uses median and IQR to match the box plot statistics.

```{r hh-composition, fig.width = 4}

hh_summary = exposure_data %>%
  filter(HH > 0) %>%
  mutate(   
    S_pctHH = if_else(HH == 0, 0, 100 * (10^S/10^HH)),
    M_pctHH = if_else(HH == 0, 0, 100 * (10^M/10^HH))
    ) %>%
  group_by(virus, FamilyID, obs_infected) %>%
  summarise_at(vars(S_pctHH, M_pctHH), funs(mean = mean, median = median))

hh_summary %>%
  ungroup() %>%
  select(-obs_infected) %>%
  gather(stat, est, -virus, -FamilyID) %>%
  group_by(virus, stat) %>%
  summarize_if(is.double, funs(mean, median, IQR = IQR_range_str, range = range_str)) %>%
  mutate_if(is.double, round, digits = 2) %>%
  filter(str_detect(stat, "mean")) %>%
  mutate(
    stat = substr(stat, 1, 1)
  ) %>%
  rename(`Household member` = stat) %>%
  rename_at(vars(-IQR), funs(str_to_title)) %>%
  kable(caption = "Percent household composition") %>% kable_styling(full_width = F)
  
hh_summary %>%
  ggplot(aes(x = virus, y = S_pctHH_mean)) +
  geom_boxplot() +
  geom_point(aes(colour = factor(obs_infected))) +
  xlab("") +
  scale_colour_manual("", values = infection_labels$colours, breaks = infection_labels$breaks,
                      labels = infection_labels$labels) +
  ylab("Percent of household shedding attributable to secondary children") 

```


## Sensitivity analysis

In the sensitivity analysis, we'd like to assess whether the interpolation of exposures affect the results.

```{r interpolate-plot}

exposure_data_summary %>%
  ungroup() %>%
  mutate(
    idpar = fct_recode(fct_rev(idpar), 
                       "Secondary\nChildren" = "S", "Mother" = "M", "Household\nSum" = "HH")
    ) %>%
  ggplot(aes(x = idpar, y = interpolated_pct)) +
  geom_boxplot() +
  geom_point() +
  xlab("") +
  ylab("Percent of weekly exposures interpolated") +
  facet_wrap(~virus)

```

```{r interpolate-exposure}

exposure_data_summary %>% 
  ungroup() %>%
  subset(idpar != "HH") %>%
  gather(stat, estimate, mean, maximum, pct_pos) %>%
  mutate(
    idpar = fct_recode(fct_rev(idpar), 
                       "Secondary\nChildren" = "S", "Mother" = "M")
    ) %>%
  ggplot(aes(x = interpolated_pct, y = estimate, colour = factor(obs_infected))) +
  geom_point() +
  ylab("") +
  xlab("Percent of weekly exposures interpolated") +
  facet_grid(stat ~ virus+idpar, scales = "free_y", switch = "y") +
  theme(strip.placement = "outside")

```

```{r}
interp_limit = c(25, 50) 

map_df(interp_limit, function(i){
  exposure_data_summary %>% 
    ungroup() %>%
    subset(idpar != "HH" & interpolated_pct <= i) %>% 
    gather(stat, estimate, mean, maximum, pct_pos) %>%
    group_by(virus, stat, idpar, obs_infected) %>%
    summarize(
      median = median(estimate)
    ) %>%
    mutate(obs_infected = recode_factor(obs_infected, `1` = "Transmission", `0` = "No transmission", .ordered = T)) %>%
    spread(obs_infected, median) %>%
    mutate(
      max_interpolation = paste0(i, "%"),
      Difference = `Transmission`-`No transmission`) %>%
    select(virus, stat, idpar, max_interpolation, Difference)

}) %>%
  bind_rows(select(mutate(subset(overall_summary, idpar != "HH"), max_interpolation = "All"), 
                   virus, stat, idpar, max_interpolation, Difference)) %>%
  spread(max_interpolation, Difference) %>%
  arrange(virus, stat) %>%
  kable(digits = 2) %>%
  kable_styling() %>%
  collapse_rows(1:2)


```
