---
title: "Functions used in the analysis"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output:
  workflowr::wflow_html:
    code_folding: none
editor_options:
  chunk_output_type: console
---


## Processing functions (`processing_functions.R`)

These functions were used for data manipulation generally related to presentation in tables and figures.

```{r, code=readLines("../code/processing_functions.R"), eval = F}
```

Labels for exposure analysis plots:

```{r, code=readLines("../code/plot_labels.R"), eval = F}
```

## Model fitting code

###  Model data preparation (`risk_fit_functions.R`)

These are functions that prepare the data for model fitting.

```{r, code=readLines("../code/risk_fit_functions.R"), eval = F}
```

### Optimization functions (`optim_functions.R`)

These are wrappers around the optim function to fit the model parameters. The tidy functions process the model output into a tidy results format.

```{r, code=readLines("../code/optim_functions.R"), eval = F}
```

## Build project (`build_all.R`)

Builds the project website by compiling all of the Rmarkdown files using the workflowr API.

```{r, code=readLines("../code/build_all.R"), eval = F}
```
