---
title: "Transmission risk estimation"
author: "Bryan Mayer"
date: "2019-05-07"
output: workflowr::wflow_html
---

## Setup

```{r,  warning = F, message = F, echo = F}
knitr::opts_chunk$set(
  comment = NA,
  fig.align = "center",
  tidy = FALSE
)
```

```{r, message = F, warning = F}
library(tidyverse)
library(conflicted)
library(kableExtra)
library(cowplot)
conflict_prefer("filter", "dplyr")
```


```{r setup}
theme_set(
  theme_bw() +
    theme(panel.grid.minor = element_blank(),
          legend.position = "top")
)
exposure_data = read_csv("data/exposure_data.csv")
exposure_data_long = read_csv("data/exposure_data_long.csv") %>%
  mutate(exposure = if_else(count == 1, 0, 10^(count)))

# duplicating the data to do sensitivity analysis for infected vs uninfected
model_data = exposure_data %>%
  subset(obs_infected == 1) %>%
  mutate(cohort = "Infected only") %>%
  bind_rows(mutate(exposure_data, cohort = "All"))

model_data_long = exposure_data_long %>%
  subset(obs_infected == 1) %>%
  mutate(cohort = "Infected only") %>%
  bind_rows(mutate(exposure_data_long, cohort = "All"))

model_formulas = paste("infectious_rev", 
                       c("M_raw", "S_raw", "HH_raw"),
                       sep = " ~ ")

```

## Model setup

The probability of being uninfected after a single, weekly exposure can written as following:

$$ s(i) = exp(-\beta_0  - \beta_{E}E_i) $$
or
$$ s(i) = exp(-\lambda(i)) $$

Instead of continuous time, we consider discretized time denoted by $i \in \{1, ..., n+1\}$ for n+1 weeks of exposures (with n survived exposures). The likelihood follows for a single, infected participant (note that in discrete time we don't apply an instantaneous hazard but use 1 - S(n) for the infectious week):

$$ L_j(n_j+1) = \prod_{i = 1}^{n_j} s_j(i) * (1-s_j(n_j+1))$$

for the j$^{th}$ participant with a unique set of total exposures. We next setup the log-likelihood for the population (m participants) and use $\Delta$ to denote an observed infection.

$$ \sum_{j = 1}^{m} log L_j(n_j) = \sum_{j = 1}^{m}\sum_{i = 1}^{n_j} log(s_j(i)) +  \sum_{j = 1}^{m} \Delta_j * log(1-s_j(n_j+1))$$


The following assumptions are used: at-risk individuals are independent and the risk associated with a weekly exposures is unique from other exposures (i.e., non-infectious exposure weeks are exchangeable). From this formulation, we can find the maximum likelihood estimators for the parameters by minimizing the negative log likelihood. Both parameters are solved numerically.

The null model has the following form for weekly risk:

$$ s_0(i) = exp(-\beta_0) $$

I.e., risk is constant and not affected by household exposure.  The null model has a simplified log-likelihood 

$$ \sum_{j = 1}^{m} log L_j(n_j) = -\beta_0 \sum_{j = 1}^{m}n_j + log(1-exp(-\beta_0)) \sum_{j = 1}^{m} \Delta_j $$

with closed-form solution for the null risk 

$$\hat{\beta_0} = log(1 + \frac{\sum_{j = 1}^{m} \Delta_j}{\sum_{j = 1}^{m}n_j})$$

## Run models

```{r null-model}

# note negative loglik is calculated
null_risk_mod = model_data %>% 
  group_by(virus, FamilyID, cohort) %>%
  summarize(
    infected = max(infectious_1wk),
    surv_weeks = max(c(0, infant_wks[which(infectious_1wk == 0)]))
  ) %>%
  group_by(virus, cohort) %>%
  summarize(
    null_beta = log(1+sum(infected)/sum(surv_weeks)),
    null_loglik = null_beta * sum(surv_weeks) - log(1-exp(-null_beta)) * sum(infected)
  )

```

```{r fitted-models}
surv_logLikE = function(x0, dat) surv_logLik(c(-Inf, x0), dat)

risk_mod = model_data_long %>%
  group_by(virus, idpar, cohort) %>%
  nest() %>%
  mutate(
    total = map_dbl(data, ~n_distinct(.x$FamilyID)), 
    total_infected = map_dbl(data, ~sum(.x$infectious_1wk)),
    fit_mod = map(data,  ~optim(c(-12, -12), surv_logLik, dat = .x,  
                                method = "L-BFGS-B")),
    fit_modE = map(data, ~optimize(surv_logLikE, c(-20, 0), dat = .x)),
    fit_res = map(fit_mod, tidy_fits)
  ) %>%
  left_join(null_risk_mod, by = c("virus", "cohort")) %>%
  unnest(fit_res) %>%
  mutate(
    null_betaE = map_dbl(fit_modE, ~exp(`$`(.x, "minimum"))),
    betaE_loglik = map_dbl(fit_modE, ~`$`(.x, "objective")),
    LLR_stat = 2 * (null_loglik - loglik),
    LLR_statE = 2 * (betaE_loglik - loglik),
    pvalue = pchisq(LLR_stat, 1, lower.tail = F),
    pvalue_beta0 = pchisq(LLR_statE, 1, lower.tail = F)
  )

```

## Model results

Outline:

- Household and secondary children estimates were very similar: household risk coefficients were within 3% of secondard children (usually < 1%). So results will focus just on mother and children estimates. 

```{r S-vs-HH}

risk_mod %>% 
  subset(idpar %in% c("HH", "S")) %>%
  select(cohort, virus, idpar, beta0, betaE) %>%
  gather(parameter, value, beta0, betaE) %>%
  unite(temp, idpar, parameter) %>%
  group_by(cohort, virus) %>%
  spread(temp, value) %>%
  select(cohort, virus, S_beta0, HH_beta0,S_betaE, HH_betaE) %>%
  mutate(pct_S0 = 100 * (1- HH_beta0/S_beta0),
         pct_SE = 100*(1-HH_betaE/S_betaE)) %>%
  kable(caption = "HH estimates were within 1.5% of S estimates for full cohort.", digits = 2)

```


```{r diagnostic}

family_risk = model_data_long %>%
  subset(idpar != "HH") %>%
  group_by(virus, FamilyID, idpar, cohort, obs_infected) %>%
  summarize(
    total_weeks = max(infant_wks),
    total_exposure = sum(exposure)
  ) %>%
  left_join(subset(risk_mod), by = c("virus", "idpar", "cohort")) %>%
  mutate(
    pr_survival = exp(-beta0 * total_weeks - betaE * total_exposure)
  ) %>%
  select(-contains("lik"), -data, -fit_mod, -contains("null"), -pvalue, -LLR_stat)

family_risk %>%
ggplot(aes(x = factor(obs_infected), y = pr_survival, colour = idpar)) +
  geom_boxplot() +
  geom_point(position = position_dodge(width = 0.75)) +
  facet_grid(virus ~ cohort, scales = "free_x")

family_risk %>%
  subset(cohort == "All") %>%
  ggplot(aes(x = factor(obs_infected), y = total_exposure, colour = idpar)) +
  geom_boxplot() +
  geom_point(position = position_dodge(width = 0.75)) +
  scale_y_log10() +
  facet_grid(virus ~ cohort, scales = "free_x")

```

```{r plot-coef}
risk_mod %>%
 ggplot(aes(x = idpar, y = betaE, colour = cohort, shape = pvalue < 0.05)) +
  geom_point(size = 2, position = position_dodge(width = 0.5)) +
  scale_y_log10("Estimated weekly exposure risk coefficient") +
  scale_x_discrete("Exposure source", drop = F) +
  facet_wrap(~ virus) +
  scale_shape_manual("Sig.", values = c(1,16)) +
  theme(legend.position = "top")
```


```{r prediction}

risk_data = exposure_data_long %>%
  mutate(
    exposure_cat = floor(count) + 0.5
  ) %>%
  group_by(idpar, virus, exposure_cat) %>%
  summarize(
    total_infected = sum(infectious_1wk),
    risk = mean(infectious_1wk)
  )

test_doses = 0:14/2

pred_fun = function(mod, pred){
  exposure_max = log10(max(mod$data[, pred]))

  pred_dat = tibble(exposure = 10^(seq(0, exposure_max, length = 100)))
  pred_dat[, pred] = pred_dat$exposure
  pred_dat %>%
    mutate(
      risk = 1 - predict(mod, pred_dat, type = "response"),
    ) %>%
    select(exposure, risk)
}

risk_prediction = risk_mod %>%
  #subset(cohort == "All") %>%
  group_by(cohort, virus, idpar) %>%
  nest() %>%
  mutate(
    pred_res = map(data, ~with(.x, tibble(
      exposure = seq(0, 6.75, length = 100),
      risk_full = 1 - exp(-beta0 - betaE * 10^exposure),
      riskE = 1 - exp(-null_betaE * 10^exposure)
    )))
  ) %>%
  unnest(pred_res) %>%
  gather(risk_est, risk, risk_full, riskE)

risk_data  %>%
  ggplot(aes(x = exposure_cat, y = risk)) +
  geom_bar(stat = "identity") +
  geom_line(data = risk_prediction, aes(x = exposure, colour = cohort, linetype = risk_est)) +
  scale_y_continuous("Estimated average risk", breaks = 0:4/4) +
  scale_x_continuous("Log10 Exposure viral load") +
  facet_grid(virus~idpar) +
  theme(legend.position = "top")

```
