---
title: "Model fitting and sensitivity analysis"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Background

This R markdown script performs all of the model fitting and sensitivity analysis. For pre-processing of the model data and a brief background on the model, see [transmission model data setup and background](setup-model-data.html). Details on the sensitivity analysis are explained in the Appendix of the manuscript.

# Setup

This document was compiled after models have been previously fit (because they take several minutes to run). To refit the models, set any of the following flags (in the setup chunk) to TRUE: `fit_loo` (leave-one-out), `fit_sens` (interpolation sensitivity), or `fit_final`. Fitted parameters shouldn't change unless the seed is changed (in model-fitter chunk), as the seed determined the LHS-sampled initial values.

```{r,  warning = F, message = F, echo = F}
knitr::opts_chunk$set(
  comment = NA,
  fig.align = "center",
  tidy = FALSE
)
```

```{r, message = F, warning = F}
library(tidyverse)
library(conflicted)
library(kableExtra)
library(cowplot)
library(scales)
library(lemon)
library(lhs)
theme_set(
  theme_bw() +
    theme(panel.grid.minor = element_blank(),
          legend.position = "top")
)

conflict_prefer("filter", "dplyr")
```

```{r setup}

source("code/risk_fit_functions.R")
source("code/processing_functions.R")
source("code/optim_functions.R")
load("output/preprocess-model-data/model_data.RData")

raw_dat = read_csv("data/PHICS_transmission_data.csv") %>%
    subset(idpar != "P" & !virus %in% c("HHV-8", "HSV", "EBV") & 
           !(FamilyID == "AZ" & virus == "HHV-6"))

fit_loo = F
fit_sens = F
fit_final = F
if(!fit_loo) {
  loo_sensitivity = read_rds("output/sensitivity-analysis/loo_fits.rds")
  loo_sensitivity_final = read_rds("output/sensitivity-analysis/loo_fits_final.rds")
}
if(!fit_sens) load("output/sensitivity-analysis/sens_fits.RData")


max_consecutive = function(x, match_var = T){
  if(!any(match_var %in% x)) return(0)
  max(rle(x)$lengths[rle(x)$values == match_var])
}

sci_not <- function (x) ifelse(log10(x) <= -8, list(bquote(phantom(0) <= 10^-8)),
                         trans_format("log10",math_format(10^.x))(x))

```


```{r model-fitter}

# use lhs initials
set.seed(10)
marginal_initials = -(20 * optimumLHS(25, 2))
combined_initials = -(20 * optimumLHS(25, 3))

# fits null, individual, and combined models
# assumed the initial valued are gliobally defined (above)
fit_models_tidy = function(wide_mod_dat, long_mod_dat){
     
  null_risk_mod = wide_mod_dat %>% 
    group_by(virus, FamilyID) %>%
    summarize(infected = max(infectious_1wk),
              surv_weeks = max(c(0, infant_wks[which(infectious_1wk == 0)]))) %>%
    group_by(virus) %>%
    summarize(
      null_beta = -log(1 - sum(infected) / sum(surv_weeks)),
      null_loglik = null_beta * sum(surv_weeks) - log(1 - exp(-null_beta)) * sum(infected)
    )
    
  individual_mod = long_mod_dat %>%
    group_by(virus, idpar) %>%
    nest() %>%
    mutate(
      likdat = map(data, create_likdat),
      total = map_dbl(data, ~ n_distinct(.x$FamilyID)),
      fit_mod = map(likdat, marginal_fitter, initials = marginal_initials),
      fit_res = map(fit_mod, tidy_fits)
    ) %>%
    unnest(fit_res) %>%
    left_join(null_risk_mod, by = "virus") %>%
    mutate(
      LLR_stat = 2 * (null_loglik - loglik),
      pvalue = pchisq(LLR_stat, 1, lower.tail = F)
    ) %>%
    select(-data,-likdat,-fit_mod) %>%
    mutate(model = "Individual")
  
  combined_mod = wide_mod_dat %>%
    group_by(virus) %>%
    nest() %>%
    mutate(
      likdat = map(data, create_likdat_combined),
      total = map_dbl(data, ~ n_distinct(.x$FamilyID)),
      fit_mod = map(likdat, combined_fitter, initials = combined_initials),
      fit_res = map(fit_mod, tidy_fits_combined)
    ) %>%
    unnest(fit_res) %>%
    left_join(null_risk_mod, by = "virus") %>%
    mutate(
      LLR_stat = 2 * (null_loglik - loglik),
      pvalue = pchisq(LLR_stat, 2, lower.tail = F)
    ) %>%
    select(-data,-likdat,-fit_mod) %>%
    gather(idpar, betaE, betaM, betaS) %>%
    mutate(model = "Combined", idpar = str_remove_all(idpar, "beta"))
  
  bind_rows(individual_mod, combined_mod)
  
}


```

# Leave-one-out model fitting

## All data

```{r loo-mod-fits, eval = fit_loo}

loo_sensitivity = model_data %>%
  select(virus, FamilyID) %>%
  distinct() %>%
  bind_rows(crossing(virus = c("CMV", "HHV-6"), FamilyID = "All")) %>%
  pmap_df( ~ (
    fit_models_tidy(
      wide_mod_dat = subset(model_data, virus == ..1 & FamilyID != ..2),
      long_mod_dat = subset(model_data_long, virus == ..1 &
                              FamilyID != ..2 & idpar != "HH")
    ) %>%
      mutate(FamilyID = ..2)
  ))%>%
  ungroup()
 
write_rds(loo_sensitivity, "output/sensitivity-analysis/loo_fits.rds")

```


```{r loo-all, fig.height=7}

loo_pl = loo_sensitivity %>%
  mutate(
    virus = factor(virus, level = c("HHV-6", "CMV")),
    FamilyID2 = factor(
      FamilyID,
      levels = c(unique(model_data$FamilyID),
                 "All"),
      labels = c(unique(model_data$FamilyID),
                 "All data")
    )
  ) %>% 
  ggplot(aes(x = FamilyID, y = pmax(betaE, 1e-8), color = idpar)) +
  geom_point(size = 2) +
  geom_path(aes(group = idpar)) +
  scale_y_log10("Est. exposure risk coef.") +
  scale_x_discrete("FamilyID left out") +
  facet_grid(virus~model) +
  scale_color_manual("Exposure Source", values = c("#F8766D", "grey30"),
                     breaks = c("S", "M"))+
  theme(legend.position = "top", legend.box = "vertical") +
  coord_flip()

loo_pl

```



```{r fam-ab-pl}
famAb_pl = raw_dat %>%
  subset(FamilyID == "AB" & virus == "HHV-6") %>%
  mutate(idpar2 = str_remove_all(PatientID, "AB-")) %>%
  ungroup() %>%
  select(infant_days, idpar2, count) %>%
  pivot_wider(names_from = idpar2, values_from = count) %>%
  mutate(S = if_else(is.na(S2), S1, log10(10^(S1) + 10^(S2)))) %>%
  gather(idpar2, count, M, S, S1, S2) %>%
  subset(!is.na(count)) %>%
  mutate(
    idpar2 = factor(
      idpar2,
      levels = c("M", "S1", "S2", "S"),
      labels = c("Mother", "Secondary child 1", "Secondary child 2",
                 "Secondary children (sum)")
    )    
  ) %>%
  ggplot(aes(x = infant_days, y = count, colour = idpar2,
             shape = idpar2)) +
  geom_point() +
  scale_x_continuous("Days after infant AB birth") +
  scale_y_continuous(expression(paste('HHV-6 log'[10], " VL"))) +
  scale_color_manual("Household AB", values = c("#F8766D", "blue", "green", "grey30")) +
  scale_shape_manual("Household AB", values = c(1,1,1,2)) +
  geom_line() +
  theme(legend.position = c(0.8, 0.25), 
        legend.background = element_rect(fill = NA))
famAb_pl
```


```{r loo-mod-final-fits, eval = fit_loo}

loo_sensitivity_final = model_data %>% 
  subset(!(FamilyID == "AB" & virus == "HHV-6")) %>%
  select(virus, FamilyID) %>%
  distinct() %>%
  bind_rows(crossing(virus = c("CMV", "HHV-6"), FamilyID = "All")) %>%
  pmap_df( ~ (
    fit_models_tidy(
      wide_mod_dat = subset(model_data, virus == ..1 & FamilyID != ..2 &
                              !(FamilyID == "AB" & virus == "HHV-6")),
      long_mod_dat = subset(model_data_long, virus == ..1 &
                              FamilyID != ..2 & idpar != "HH" &
                              !(FamilyID == "AB" & virus == "HHV-6"))
    ) %>%
      mutate(FamilyID = ..2)
  )) %>%
  ungroup()
  
write_rds(loo_sensitivity_final, "output/sensitivity-analysis/loo_fits_final.rds")

```

## Remove FamilyID AB from HHV-6

```{r loo-final, fig.height=7}

loo_pl2 = loo_sensitivity_final %>%
  mutate(
    virus = factor(virus, level = c("HHV-6", "CMV")),
    FamilyID2 = factor(
      FamilyID,
      levels = c(unique(model_data$FamilyID),
                 "All"),
      labels = c(unique(model_data$FamilyID),
                 "All data")
    )
  ) %>% 
  ggplot(aes(x = FamilyID2, y = pmax(betaE, 1e-8), color = idpar)) +
  geom_point(size = 2) +
  geom_path(aes(group = idpar)) +
  scale_y_log10("Est. exposure risk coef.") +
  scale_x_discrete("FamilyID left out") +
  facet_grid(virus~model) +
  scale_color_manual("Exposure Source", values = rev(c("grey30", "#F8766D")),
                     breaks = c("S", "M")) +
  theme(legend.position = "top", legend.box = "vertical") +
  coord_flip()

loo_pl2

```

# Interpolation and infected only

```{r interp-data}

# interpolation info by family
interpolated_summary = model_data_long %>%
  subset(idpar != "HH" & !(FamilyID == "AB" & virus == "HHV-6")) %>%
  group_by(virus, FamilyID, obs_infected, idpar) %>%
  arrange(virus, FamilyID, obs_infected, idpar, infant_wks) %>%
  summarize(
    consecutive_int = max_consecutive(interpolated),
    interpolated_pct = 100*mean(interpolated)
    ) %>%
  select(FamilyID, virus, idpar, obs_infected, interpolated_pct, consecutive_int) %>%
  group_by(FamilyID, virus) %>%
  mutate(
    max_interp = max(interpolated_pct),
    max_consecutive_interp =  max(consecutive_int)
    )

```

## Summarizing interpolation

### Total % interpolation:

```{r sensitivity-dat-summary}

interpolated_summary %>%
  group_by(virus) %>%
  select(-consecutive_int, -max_consecutive_interp) %>%
  spread(idpar, interpolated_pct) %>%
  summarize(
    all_M_lte_S = all(M <= S),
    max_mom = max(M),
    max_sec_child = max(S),
    max_household = max(max_interp)
  ) %>%
  mutate_if(is.numeric, round) %>%
  kable(caption = "max percent interpolated by exposure (mothers always have less interpolation)") %>%
    kable_styling(full_width = F)

interpolated_summary %>%
  select(FamilyID, virus, obs_infected, max_interp) %>%
  distinct() %>%
  group_by(virus, obs_infected) %>%
  summarize(
    n = n(),
    none = sum(max_interp == 0),
    lt20 = sum(max_interp < 20),
    gte20 = sum(max_interp >= 20)
  ) %>%
  kable(caption = "Household percent interpolated") %>%
    kable_styling(full_width = F)

interpolated_summary %>%
  bind_rows(mutate(ungroup(interpolated_summary), obs_infected = 2)) %>%
  group_by(virus, idpar, interpolated_pct, obs_infected) %>%
  summarize(total_households = n()) %>%
  group_by(virus, idpar, obs_infected) %>%
  arrange(virus, idpar, obs_infected, interpolated_pct) %>%
  mutate(cumulative_households = cumsum(total_households)) %>%
  ungroup() %>%
  mutate(households = factor(obs_infected, levels = 0:2,
                               labels = c("Uninfected", "Infected", "Total")),
         idpar = factor(idpar, levels = c("S", "M"),
                   labels = c("Secondary children", "Mother"))) %>%
  ggplot(aes(x = interpolated_pct, y = cumulative_households, colour = households)) + 
  geom_step() +
  scale_x_continuous("Total interpolation (%)") +
  scale_y_continuous("Total households", breaks = 0:6*5) +
  labs(color = "Household infection status") +
  facet_grid(virus~idpar)

```

### Summarizing interpolation

There are several households that have at least one long stretch of interpolated values. For both viruses, consecutive interpolation generally correlates with total interpolation. For HHV-6, there were two households with lots of scattered (non-consecutive) interpolations. The mother with 50% interpolation is for an infant infected on week 2 where the first week was back interpolated.

```{r}

interpolated_summary %>%
  group_by(virus) %>%
  select(-interpolated_pct, -max_interp) %>%
  spread(idpar, consecutive_int) %>%
  group_by(virus) %>%
  summarize(
    all_M_lte_S = all(M <= S),
    max_mom = max(M),
    max_sec_child = max(S),
    max_household = max(max_consecutive_interp)
  ) %>%
  mutate_if(is.numeric, round) %>%
  kable(caption = "max consecutive interpolated weeks by exposure (mothers always have less interpolation)") %>%
    kable_styling(full_width = F)

```


```{r interp-pl}
interp_pl = interpolated_summary %>%
  ungroup() %>%
  mutate(
    virus = factor(virus, 
                   levels = c("HHV-6",  "CMV")),
    idpar = factor(
      idpar,
      levels = c("S", "M"),
      labels = c("Secondary children", "Mother")
    )
  ) %>%
  ggplot(aes(x = interpolated_pct, y = consecutive_int, colour = idpar)) +
  geom_point() +
  scale_color_manual("Exposure Source", values = c("grey30", "#F8766D")) +
  scale_y_continuous("Max consecutive interp. (wks)", breaks = 0:12) +
  scale_x_continuous("Interp. exposures (%)", breaks = 0:6*10) +
  facet_grid(virus ~ idpar) 

```


## Sensitivity analysis by interpolation and infected status

To assess how families with varying interpolated exposures affect the estimate, we refit the models restricting maximum % interpolation. For the marginal models, we look at this by exposure source. For combined model, we look at total household interpolation (driven by secondary children missing data). Any rules regarding interpolation-based exclusion would be applied at the household-level for consistency. 

### Set up sensitivity data

Create datasets with varying allowed total interpolation. Because interpolation, this is more complicated than using the wrapper above.

```{r sens-data}

# the set of unique interpolated values for sensitivity analysis
interpolation_idpar_pct = interpolated_summary %>%
  ungroup() %>%
  select(virus, max_interp, idpar, interpolated_pct, FamilyID) %>%
  mutate_if(is.numeric, round) %>%
  group_by(virus, max_interp, interpolated_pct, idpar) %>%
  summarize(total = n_distinct(FamilyID)) %>%
  ungroup() %>%
  select(virus, idpar, interpolated_pct) %>%
  distinct() 

interpolation_max_pct = interpolated_summary %>%
  ungroup() %>%
  select(virus, max_interp) %>%
  mutate(max_interp = round(max_interp)) %>%
  distinct()

# this is used for the marginal models
# left join used to subset
sensitivity_data_long = interpolation_idpar_pct %>%
  subset(interpolated_pct >= 20) %>%
  pmap_df(~(
     interpolated_summary %>%
       subset(virus == ..1 & idpar == ..2 & interpolated_pct <= ..3) %>%
       ungroup() %>% 
       select(FamilyID, idpar, virus, interpolated_pct) %>%
       left_join(model_data_long, by = c("virus", "FamilyID", "idpar")) %>%
       mutate(cohort = as.character(..3))
     )) 

# left_join used to subset
sensitivity_data = interpolation_max_pct %>%
  subset(max_interp >= 20) %>%
  pmap_df(~(interpolated_summary %>%
              ungroup() %>%
              select(FamilyID, max_interp, virus) %>%
              distinct() %>%
              subset(max_interp <= ..2 & virus == ..1) %>%
              select(FamilyID, virus, max_interp) %>%
              left_join(model_data, by = c("virus", "FamilyID")) %>%
              mutate(cohort = as.character(..2))
           ))

```

### Run Models

```{r sensitivity-models, eval = fit_sens}

marginal_sensitivity = map_df(c("All", "Infected only"), function(i){
  if(i == "Infected only") sensitivity_data_long = subset(sensitivity_data_long, obs_infected == 1)
  
  sensitivity_data_long %>%
    group_by(virus, idpar, cohort) %>%
    nest() %>%
    mutate(
      data_cohort = i,
      likdat = map(data, create_likdat),
      total = map_dbl(data, ~ n_distinct(.x$FamilyID)),
      total_infected = map_dbl(data, ~ sum(.x$infectious_1wk)),
      fit_mod = map(likdat, marginal_fitter, initials = marginal_initials),
      fit_res = map(fit_mod, tidy_fits)
    ) %>%
    unnest(fit_res)  %>%
    select(-data, -likdat, -fit_mod)
  }) %>%
  mutate(model = "Individual")

combined_sensitivity = map_df(c("All", "Infected only"), function(i){
  if(i == "Infected only") sensitivity_data = subset(sensitivity_data, obs_infected == 1)

  sensitivity_data %>%
    group_by(virus, cohort) %>%
    nest() %>%
    mutate(
      data_cohort = i,
      likdat = map(data, create_likdat_combined),
      total = map_dbl(data, ~ n_distinct(.x$FamilyID)),
      total_infected = map_dbl(data, ~ sum(.x$infectious_1wk)),
      fit_mod = map(likdat, combined_fitter, initials = combined_initials),
      fit_res = map(fit_mod, tidy_fits_combined)
    ) %>%
    unnest(fit_res)  %>%
    select(-data, -likdat, -fit_mod)
  }) %>%
  gather(idpar, betaE, betaM, betaS) %>%
  mutate(model = "Combined", idpar = str_remove_all(idpar, "beta"))

save(marginal_sensitivity, combined_sensitivity, 
     file = "output/sensitivity-analysis/sens_fits.RData")

```

```{r tests}

null_risk_sens = sensitivity_data_long %>% 
  group_by(cohort, idpar, virus, FamilyID) %>%
  summarize(
    infected = max(infectious_1wk),
    surv_weeks = max(c(0, infant_wks[which(infectious_1wk == 0)]))
  ) %>%
  group_by(cohort, idpar, virus) %>%
  summarize(
    frac_infected = paste(sum(infected), n(), sep = "/"),
    null_beta = -log(1-sum(infected)/sum(surv_weeks)),
    null_loglik = null_beta * sum(surv_weeks) - log(1-exp(-null_beta)) * sum(infected)
  )

tests = marginal_sensitivity %>%
  left_join(null_risk_sens, by = c("virus", "cohort", "idpar")) %>%
  mutate(
    LLR_stat = 2 * (null_loglik - loglik),
    pvalue = pchisq(LLR_stat, 1, lower.tail = F)
  ) %>%
  arrange(virus) 

```

```{r sensitivity-results}

sens_pl = combined_sensitivity %>%
  bind_rows(marginal_sensitivity) %>%
  ungroup() %>%
  mutate(
    virus = factor(virus, 
                   levels = c("HHV-6",  "CMV")),
    idpar = factor(
      idpar,
      levels = c("S", "M"),
      labels = c("Secondary children", "Mother")
    ),
    model = factor(model, levels = c("Individual", "Combined")),
    cohort = as.numeric(cohort),
    frac_infected = paste(total_infected, total, sep = "/")
  ) %>% 
  arrange(idpar, virus, cohort) %>%
  ggplot(aes(x = factor(cohort, levels = unique(sort(cohort)),
                        labels =  c(head(unique(sort(cohort)), -1), "60\nAll\ndata")), 
             y = pmax(betaE, 1e-8), color = idpar, 
             linetype = model, shape = model)) +
  geom_point(size = 2) +
  geom_path(aes(group = interaction(idpar, model))) +
  scale_y_log10(paste("Est. exposure risk coef."), labels = sci_not) +
  scale_x_discrete("Max interpolation (%) by household") +
  facet_grid(virus ~ data_cohort) +
  scale_color_manual("Exposure Source", values = c("grey30", "#F8766D")) +
  scale_shape_discrete("Model") +
  scale_linetype_discrete("Model") +
  theme(legend.position = "top", legend.box = "vertical")
sens_pl

```


# Sensitivity plots

```{r loo-pl-final}

loo_pl_final = loo_sensitivity %>%
  subset(virus == "HHV-6") %>%
  mutate(virus = "HHV-6 (All Households)") %>%
  bind_rows(mutate(loo_sensitivity_final, 
            virus = if_else(virus == "CMV", virus, "HHV-6 (Without household AB)"))) %>%
  ungroup() %>%
  mutate(
    virus = factor(virus, 
                   levels = c("HHV-6 (All Households)",  "CMV", 
                              "HHV-6 (Without household AB)")),
    FamilyID2 = factor(FamilyID,
                       levels = c(unique(model_data$FamilyID),
                                  "All"),
                       labels = c(unique(model_data$FamilyID),
                                  "All data")),
    model = factor(model, levels = c("Individual", "Combined")),
    idpar = factor(
      idpar,
      levels = c("S", "M"),
      labels = c("Secondary children", "Mother")
      )
    ) %>% 
  ggplot(aes(x = FamilyID2, y = pmax(betaE, 1e-8), color = idpar, shape = model))+
  geom_point(size = 2) +
  geom_path(aes(group = interaction(idpar, model))) +
  scale_y_log10("Est. exposure risk coef.",
                labels = sci_not) +
  scale_x_discrete("Household left out") +
  facet_wrap(~ virus, nrow = 2) +
  scale_color_manual("Exposure Source", values = c("grey30", "#F8766D")) +
  scale_shape_discrete("Model") +
  theme(legend.position = "top", legend.box = "horizontal",
        axis.text.y = element_text(size = 6.5)) +
  coord_flip() +
  theme(legend.box = "vertical",
                         legend.direction = "vertical",
                         legend.box.just = "left",
                         legend.spacing.y = unit(0, "lines"))


```

```{r sens-final-pl1, fig.height=8}
noleg = theme(legend.position = "none")

leg = get_legend(loo_pl_final +
                   theme(legend.box = "vertical",
                         legend.direction = "vertical",
                         legend.box.just = "left",
                         legend.spacing.y = unit(0, "lines")))

plot_grid(reposition_legend(loo_pl_final, position =  'center', panel = 'panel-2-2', 
                            plot = F),
          famAb_pl + theme(legend.position = "right"), nrow = 2, rel_heights = c(3,1),
          labels = c("(a)", "(b)"), vjust = c(1,-0.25))

```

```{r sens-final-pl2, fig.height=8}
leg2 = get_legend(loo_pl_final +
                   theme(legend.box = "horizontal",
                         legend.direction = "vertical",
                         legend.box.just = "left",
                         legend.spacing.y = unit(0, "lines")))

leg2 = get_legend(interp_pl)

plot_grid(
  plot_grid( interp_pl +noleg, 
             sens_pl + guides(colour = "none"), 
             labels = c("(a)", "(b)"), ncol = 1, align = "v", rel_heights = c(5,6)),
            leg2, ncol = 1, rel_heights = c(12, 1))

```


# Final model

```{r, eval = fit_final}

final_model = fit_models_tidy(
  wide_mod_dat = subset(model_data,!(FamilyID == "AB" &
                                       virus == "HHV-6")),
  long_mod_dat = subset(model_data_long,!(FamilyID == "AB" &
                                            virus == "HHV-6"))
)
  
write_rds(final_model, "output/final-model/final_model.rds")


```
