---
title: "Transmission risk estimation"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
---

This is the main analysis and results for the manuscript. For pre-processing of the model data and a brief background on the model, see [transmission model data setup and background](setup-model-data.html). For full model fitting and sensitivity analysis, see the analysis of [model fitting and sensitivity analysis](transmission-risk-sensitivity.html). The sensitivity analysis had downstream effects on the final model analysis presented here.

# Data Setup

```{r knitr-opt,  warning = F, message = F, echo = F}
knitr::opts_chunk$set(
  comment = NA,
  fig.align = "center",
  tidy = FALSE
)
```

```{r packages-data, message = F, warning = F}
library(tidyverse)
library(conflicted)
library(kableExtra)
library(cowplot)
library(scales)
library(lhs)
source("code/processing_functions.R")
conflict_prefer("filter", "dplyr")

theme_set(
  theme_bw() +
    theme(panel.grid.minor = element_blank(),
          legend.position = "top")
)
load("output/preprocess-model-data/model_data.RData")
final_model = read_rds("output/final-model/final_model.rds")

# function calculate IDX (e.g., X (prob) = 50 for ID50)
IDX_calc = function(b0, bE, prob){
  (-log(1-prob) - b0)/(bE)
}


```


# Model results

```{r res-tab}
options(knitr.kable.NA = '')

exposure_summary = model_data_long %>%
  group_by(virus, idpar) %>%
  summarize(
    mean_exposure = 10^mean(count[count > 0]),
    mean_exposure2 = 10^mean(count),
    max_exposure = max(exposure)
  ) %>%
  mutate(parameter = paste0("b", idpar)) %>%
  ungroup()
  
combined_wide = final_model %>%
  subset(model == "Combined") %>%
  pivot_wider(names_from = idpar, values_from = betaE) %>%
  rename(betaM = M, betaS = S)

final_model_tab = final_model %>%
  select(virus, model, idpar, beta0, betaE, null_beta, pvalue, loglik, null_loglik) %>%
  arrange(desc(virus),  desc(model), idpar) %>%
  gather(parameter, estimate, beta0, betaE, null_beta) %>%
  select(virus, model, idpar, parameter, estimate, pvalue, null_loglik, loglik) %>%
  mutate(
    model_cat = case_when(
      parameter == "null_beta" ~ "Constant",
      model == "Combined" ~ "CM",
      TRUE ~ idpar
    ),
    loglik = if_else(parameter == "null_beta", null_loglik, loglik),
    parameter = case_when(
      parameter == "null_beta" ~ "b0",
      parameter == "betaE" ~ str_c("b", idpar),
      TRUE ~ str_remove(parameter, "eta")
    ),
    model = factor(
      model_cat,
      levels = c("Constant", "S", "M", "HH", "CM"),
      labels = c(
        "Constant risk",
        "Secondary children",
        "Mother",
        "Household sum",
        "Combined model"
      )
    ),
    pvalue = if_else(model == "Constant risk", NA_real_, pvalue)
  ) %>%
  ungroup() %>%
  select(-model_cat, -null_loglik, -idpar) %>%
  distinct()


beta0_ests = final_model_tab %>%
  subset(model != "Constant risk" & parameter == "b0") %>%
  select(virus, model, estimate) %>%
  rename(beta0 = estimate)

```


```{r est-res-tab}

final_model_tab %>%
  select(virus, model, parameter, estimate, loglik) %>%
  arrange(desc(virus), model) %>%
  write_csv("output/results-tables/supp_table4.csv") %>%
  mutate_if(is.numeric, format, digits = 3) %>%
  kable(digits = 3, caption = "Parameter model estimates") %>%
  kable_styling(full_width = F) %>%
  collapse_rows(c(1:2, 5), valign = "top")
 
```

```{r mean-max-table}

exposure_summary %>%
  arrange(desc(virus), desc(idpar)) %>%
  select(-mean_exposure2, -parameter) %>%
  mutate_if(is.numeric, format, digits = 3, scientific = T) %>%
  kable(digits = 3, caption = "Parameter model estimates") %>%
  kable_styling(full_width = F) %>%
  collapse_rows(c(1), valign = "top")

```

```{r res-ind-risk}

constant_risk_res = final_model_tab %>%
  subset(model == "Constant risk") %>%
  mutate(
    model = as.character(model),
    constant_risk = 1 - exp(-estimate)
    ) %>%
  select(virus, model, constant_risk)

final_model_tab %>%
  subset(model %in% c("Mother", "Secondary children", "Household sum")) %>%
  pivot_wider(names_from = parameter, values_from = c(estimate)) %>%
  mutate(idpar = if_else(model == "Household sum", "HH",
                         substr(as.character(model), 1, 1))) %>%
  left_join(exposure_summary, by = c("virus", "idpar")) %>%
  mutate(
    bE = if_else(model == "Household sum", bHH, pmax(bM, bS, na.rm = T)),
    constant_risk = 1 - exp(-b0),
    exposure_risk_mean = 1 - exp(-b0 - bE * mean_exposure),
    exposure_risk_max = 1 - exp(-b0 - bE * max_exposure)
  ) %>%
  select(virus, idpar, constant_risk, mean_exposure, exposure_risk_mean, max_exposure, exposure_risk_max, pvalue) %>%
  rename(model = idpar) %>%
  bind_rows(constant_risk_res) %>%
  mutate_at(vars(mean_exposure, max_exposure), (function(x) round_away_0(log10(x), 2, trailing_zeros = T))) %>%
  mutate_at(vars(contains("risk")), (function(x) round_away_0(100*x, 2, trailing_zeros = T))) %>%
  mutate(
    pvalue = clean_pvalues(pvalue, sig_alpha = 0),
    model = factor(
      model,
      levels = c("Constant risk", "S", "M", "HH"),
      labels = c("Null", "Secondary children",
                 "Mother",
                 "Household sum")
    )
  ) %>%
  arrange(desc(virus), model) %>%
  write_csv("output/results-tables/individual_risk_tab.csv") %>%
  kable(digits = 3, caption = "Parameter model estimates") %>%
  kable_styling(full_width = F) %>%
  collapse_rows(c(1), valign = "top")
  
```


```{r combined-risk}
final_model_tab %>%
  subset(model %in% c("Combined model")) %>%
  mutate(idpar = substr(parameter, 2, 2)) %>%
  select(-parameter) %>%
  left_join(exposure_summary, by = c("virus", "idpar")) %>%
  group_by(virus, pvalue, idpar) %>%
  summarize(
    risk_comp_mean = if_else(idpar == "0", estimate, estimate * mean_exposure),
    risk_comp_max = if_else(idpar == "0", estimate, estimate * max_exposure)
  ) %>%
  ungroup() %>%
  pivot_wider(names_from = idpar, values_from = c(risk_comp_mean, risk_comp_max)) %>%
  mutate(
    constant_risk = 1 - exp(-risk_comp_mean_0)
  ) %>%
  pivot_longer(cols = c(risk_comp_mean_M, risk_comp_mean_S, 
                        risk_comp_max_M, risk_comp_max_S)) %>%
  mutate(
    risk = 1 - exp(-risk_comp_mean_0 - value),
    parm = str_remove_all(name, "risk_comp_"),
    pvalue = clean_pvalues(pvalue, sig_alpha = 0)
    ) %>%
  select(-name, -contains("risk_comp"), -value) %>%
  spread(parm, risk) %>%
  mutate_if(is.numeric, (function(x) round_away_0(100*x, 2, trailing_zeros = T))) %>%
  select(virus, constant_risk, mean_S, max_S, mean_M, max_M, pvalue) %>%
  arrange(desc(virus)) %>%
  write_csv("output/results-tables/combined_risk_tab.csv") %>%
  kable(digits = 3, caption = "Parameter risk estimates (combined model)") %>%
  kable_styling(full_width = F) %>%
  collapse_rows(c(1), valign = "top")

```

## Infectious dose calculations

```{r IDx}

final_model_tab %>%
  subset(parameter != "b0" & model != "Household sum" & virus == "HHV-6") %>%
  select(-pvalue, -loglik) %>%
  group_by(virus, model) %>%
  left_join(beta0_ests, by = c("virus", "model")) %>%
  mutate(
    exposure_source = factor(case_when(
      parameter == "bS" ~ "Secondary Children",
      parameter == "bM" ~ "Mother"
    ), levels = c("Secondary Children", "Mother")),
    ID25 = IDX_calc(beta0, estimate, 0.25),
    ID50 = IDX_calc(beta0, estimate, 0.5),
    ID75 = IDX_calc(beta0, estimate, 0.75)
  ) %>%
  ungroup() %>%
  mutate_if(is.numeric, (function(x) round_away_0(log10(x), 2, trailing_zeros = T))) %>%
  arrange(model, exposure_source) %>%
  write_csv("output/results-tables/hhv6_id50_tab.csv") %>%
  select(-virus, -estimate, -beta0, -parameter) %>%
  kable(digits = 2) %>%
  kable_styling(full_width = F) %>%
  collapse_rows(1:2, valign = "top") %>%
  add_header_above(c(" " = 2, "HHV-6 Infectious dose (ID)" = 3))

```

## Dose-response results

```{r prediction-pls}

risk_data = model_data_long %>%
  subset(idpar != "HH") %>%
  mutate(
    exposure_cat = floor(count) + 0.5
  ) %>%
  group_by(idpar, virus, exposure_cat) %>%
  summarize(
    total_exposures = n(),
    total_infected = sum(infectious_1wk),
    risk = mean(infectious_1wk)
  )

risk_grid = model_data %>%
  mutate(
    exposure_S = floor(S) + 0.5,
    exposure_M = floor(M) + 0.5
  ) %>%
  group_by(virus, exposure_S, exposure_M) %>%
  summarize(
    total_exposures = n(),
    total_infected = sum(infectious_1wk),
    risk = mean(infectious_1wk)
  )

risk_prediction_both = final_model %>%
  subset(model == "Combined") %>%
  left_join(exposure_summary, by = c("virus", "idpar")) %>%
  ungroup() %>%
  mutate(idpar = factor(idpar, levels = c("S", "M"))) %>%
  group_by(virus, idpar) %>%
  nest() %>%
  mutate(
    pred_res = map(data, ~with(.x, tibble(
      exposure = seq(0, log10(max_exposure), length = 100),
      risk = 1 - exp(-beta0 - betaE * 10^exposure)
    )))
  ) %>%
  unnest(pred_res) 


dr_plots = map(c("HHV-6", "CMV"), function(i){
  
  if(i == "CMV") {
    xlab = expression(paste("Log"[10], " CMV VL (DNA copies/mL)"))
    ybreaks = pretty_breaks()
  }
  if(i == "HHV-6") {
    xlab = expression(paste("Log"[10], " HHV-6 VL (DNA copies/mL)"))
    ybreaks = 0:4/4
  }
  
  risk_data  %>%
  subset(virus == i) %>%
    ungroup() %>%
  mutate(idpar = factor(idpar, levels = c("S", "M"))) %>%
  ggplot(aes(x = exposure_cat, y = risk)) +
  geom_bar(stat = "identity", fill = "grey50") +
  geom_line(data = subset(risk_prediction_both, virus == i),
            aes(x = exposure), size = 1.25) +
  scale_y_continuous(paste("Estimated weekly risk of", i,  "infection"), 
                      breaks = ybreaks) +
  scale_x_continuous(xlab,
                     breaks = 0:6+0.5,
                     labels = paste(0:6, 1:7, sep = "-")) +
  facet_grid(.~idpar, labeller = as_labeller(c('S' = "Secondary Children", 'M' = "Mother"))) +
  theme(legend.position = "top")
})

exposure_heats = map(c("HHV-6", "CMV"), function(i){

  risk_grid %>%
  subset(virus == i) %>%
  ggplot(aes(x = exposure_S, y = exposure_M, fill = total_exposures, 
             label = paste(total_infected, total_exposures, sep = "/"))) +
  geom_tile() +
  #geom_label(label.padding = unit(0.1, "lines"), label.size = 0, fill = "white") +
  geom_text(aes(colour = factor(total_exposures > 100)), 
            fontface = "bold", size = 3.5) +
  scale_color_manual(guide=F, values = c("white", "black")) +
  viridis::scale_fill_viridis(paste("Total", i, "exposures"), option = "viridis") +
  scale_y_continuous(expression(paste("Mother log"[10], " VL (DNA copies/mL)")), 
                     breaks = 0:6+0.5,
                     labels = paste(0:6, 1:7, sep = "-")) +
  scale_x_continuous(expression(paste("Secondary children log"[10], " VL (DNA copies/mL)")),
                     breaks = 0:6+0.5,
                     labels = paste(0:6, 1:7, sep = "-")) +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.title.x = element_text(size = 10.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        legend.key.width = unit(0.75, "cm"))

})

risk_heats = map(c("HHV-6", "CMV"), function(i){
    if(i == "CMV") {
    kbreaks = pretty_breaks()
    klabels = function(x)identity(x)
  }
  if(i == "HHV-6") {
    klabels = c(0:4/20, " >0.25")
    kbreaks = c(0:5/20)
  }
  crossing(
  virus = i,
  exposure_S = seq(0, 
                   log10(subset(exposure_summary, virus == i & idpar == "S")$max_exposure)+0.1,
                   length = 100),
  exposure_M = seq(0, 
                   log10(subset(exposure_summary, virus == i & idpar == "M")$max_exposure)+0.1,
                   length = 100)
) %>%
  left_join(combined_wide, by = "virus") %>%
  mutate(
    risk =  1 - exp(-beta0 - betaS * 10^exposure_S - betaM * 10^exposure_M)
  ) %>%
  ggplot(aes(x = exposure_S, y = exposure_M, fill = pmin(risk, 0.25))) +
  geom_tile() +
  scale_y_continuous(expression(paste("Mother log"[10], " VL (DNA copies/mL)")), 
                     breaks = 1:7) +
  scale_x_continuous(expression(paste("Secondary children log"[10], " VL (DNA copies/mL)")),
                     breaks = 1:7) +
  scale_fill_distiller(paste(i, "weekly infection prob."), 
                       palette = "RdYlBu",  breaks= kbreaks,
                        labels = klabels) +
  geom_text(data = subset(risk_grid, virus == i), aes(label = round(risk, 2)), 
            fontface = "bold", colour = "black") +
  theme_classic() +
  theme(legend.position = "bottom",
        axis.title.x = element_text(size = 10.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        legend.key.width = unit(0.68, "cm"))
})

```


```{r hhv6-dr, fig.height=8}

plot_grid(dr_plots[[1]],
  plot_grid(exposure_heats[[1]], risk_heats[[1]], nrow = 1, labels = c("(b)", "(c)"), 
            hjust = -0.2, vjust = 0.5),
  nrow = 2, labels = "(a)")

```

```{r cmv-dr, fig.height=8}

plot_grid(dr_plots[[2]],
  plot_grid(exposure_heats[[2]], risk_heats[[2]], nrow = 1, labels = c("(b)", "(c)"), 
            hjust = -0.2, vjust = 0.5),
  nrow = 2, labels = "(a)")

```
